---
title: "Estimating a p value for the difference in means between two groups"
author: "PJ Coetser"
date: "12/3/2021"
output:
  html_document:
    highlight: textmate
    css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

> **Note:** this code is a direct copy of Dr Juan Klopper youtube video ( [link](https://www.youtube.com/watch?v=XRrx95pc_Pw)) and intended for educational purposes only. Obvious and notable differences includes code written in R and added math equations.


# Resampling under the null hypothesis

## In this tutiorial

- Seeding the pseudo-random number generator for reproducible results
- Creating a Dataframe
- Sample from a set of categorical values using  xxx
- Sample from a normal distribution using xxx
- Making sub-DataFrames using the filter function (conditional selection)
- Extracting vectors from DataFrames
- Summary statistics of a continuous numerical variable using xxx
- Histogram using xxx
- Shuffle a vector
- Calculate a p-value using xxx

## Packages

```{r, message=FALSE}
library(dplyr)
library(ggplot2)
```

## Population Data

Sample Size
```{r}
n <- 1000
```

Configure pseudo-random number generator and define group probability
```{r}
set.seed(12)
group_probability <- c(0.479,0.521)
```
> Note: group probability used to mimic the number of samples included in each group, as observed in youtube

Define parameters used for data generation
```{r}
groups <- c('I','II')
mass_mean <- 100
mass_sd <- 10
```

Generate population data set
```{r}
df <- data.frame(
  ID = 1:n,
  Group = replicate(n,sample(groups,size = 1, prob = group_probability)),
  Mass = rnorm(n, mass_mean, mass_sd))

```

## Sample Data Sets

```{r}
group_I <- df |> filter(Group == 'I')
group_II <- df |> filter(Group == 'II')
```

Extract Vectors for each 

$\sum_{n=1}^{10} n^2$
