---
title: "Estimating a p value for the difference in means between two groups"
author: "PJ Coetser"
date: "12/3/2021"
output:
  html_document:
    highlight: textmate
    css: styles.css
    self_contained: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

> **Note:** This article is based on a youtube video tutorial made by [Dr Juan Klopper](https://www.linkedin.com/in/juanklopper/). In the [tutorial](https://www.youtube.com/watch?v=XRrx95pc_Pw) Juan illustrate how one can estimate a p-value for the difference in means between two groups using **Julia**.  Obvious and notable differences in this post includes code written in R.

TODO: Added math equations, additional information and update aesthetics.

## Packages

```{r, message=FALSE}
library(dplyr)
library(ggplot2)
```

## Population
Data set with $N$ members, containing a numeric variable.

```{r}
N <- 1000
```

The numeric variable (mass) are distributed normally with population mean ($\mu$) and standard deviation ($\sigma$)

```{r}
mass.mean <- 100
mass.sd <- 10
```

Generate Data Set

```{r}
set.seed(12)
population <- data.frame(Mass = rnorm(N, mass.mean, mass.sd))
```

> Note: configure pseudo-random number generator for reproducibility using seed(x)

Extract numeric variable from population as vector

```{r}
mass <- population[['Mass']]
```

## Samples

Define groups

```{r}
group <- c('I','II')
```

Randomly assign group identifier to each population member

```{r}
group_probability <- c(0.479,0.521)
population[['Group']] <- replicate(N,sample(group,size = 1, prob = group_probability))
```

> Note: group probabilities are used to mimic the number of observations included in each sample set, as observered in tutorial

Extract Sample Sets

```{r}
group.I <- population |> filter(Group == 'I')
group.II <- population |> filter(Group == 'II')
```

Extract numerical variable from sample sets as vector

```{r}
mass.I <- group.I[['Mass']]
mass.II <- group.II[['Mass']]
```

### Descriptive Statistics

Get number observations ($n$) in each sample set

```{r}
n.I <- length(mass.I)
```

$n_I = `r n.I`$

```{r}
n.II <- length(mass.II)
```

$n_{II} = `r n.II`$

Get summary stats of each sample set

```{r}
summary(mass.I)
```
```{r}
summary(mass.II)
```

### Data Visualization

Plot histogram of sample data sets overlaid on top of each other

```{r}
ggplot(population, aes(
  x=Mass, 
  fill=Group)) + 
  geom_histogram(
    position = "identity",
    alpha=0.5, 
    bins = 30)
```

### Delta between samples means

Calculate and display difference between sample means

```{r}
delta_means <- abs(mean(mass.I) - mean(mass.II))
```

$\Delta\bar{X} = `r round(delta_means,3)`$

Use random pairwise samples extracted from the population data set to generate a spectrum of possible differences between sample means

```{r}
means <- c()
resamples <- 5000

for (i in 1:resamples ) {
  mass.shuffled <- sample(mass)
  group.I.new <- mass.shuffled[1:n.I]
  group.II.new <- mass.shuffled[(n.I + 1):N]
  delta_means.new = mean(group.I.new) - mean(group.II.new)
  means <- means |> append(delta_means.new)
}
```

Plot a histogram of resampled differences between sample means

```{r}
ggplot(mapping = aes(means)) + 
  geom_histogram(
    position = "identity",
    alpha=0.5, 
    bins = 30)
```


### Probablity of difference between sample means

Manually calculate p-value of the difference between the two sample means

```{r}
p_value.manual <- (length(means[means < -delta_means]) + length(means[means > delta_means])) / resamples
```

p-value $= `r round(p_value.manual,3)`$

Use t.test() function to calculate p-value of the difference between the two sample means

```{r}
p_value.package <- t.test(mass.I, mass.II, alternative = "two.sided", var.equal = FALSE)[['p.value']]
```

p-value $= `r round(p_value.package,3)`$
